summary_stats

### 2023-07-03 scripts/13-summary_stats.Rmd
summary stats

# read in corpus
corpus <- read_parquet("/Volumes/Verbatim/v4/hansard_1998_to_2022-parquet/hansard_corpus_1998_to_2022.parquet")

@@

```{r}
corpus |> select(date, speech_no, party) |> 
  filter(!is.na(party)) |> 
  left_join(party_table_final, by=c("party" = "party_abb_hansard")) |> 
  group_by(date, party_abb_auspol, party_name_auspol) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(party_abb_auspol, party_name_auspol) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches))

# only for those parties with a party facts ID
s_by_party <- corpus |> select(date, speech_no, partyfacts_id, party) |> 
  mutate(party = as.character(ifelse(str_detect(party, "^Ind"), "IND", as.character(party)))) |> 
  group_by(date, partyfacts_id) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(partyfacts_id) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  filter(!is.na(partyfacts_id)) |> 
  left_join(party_table_final |> 
              select(-party_abb_hansard, -party_abb_auspol) |> 
              distinct() |> 
              # b/c Centre Alliance is current name and they have the same party facts ID
              filter(party_name_auspol!="Nick Xenophon Team"), 
            by="partyfacts_id")

@@

# plot
ggplot(s_by_day, aes(x=as.Date(date), y=n_distinct_speeches)) + 
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "date", y = "# distinct speeches")
```

@@

# plot
# remove 2022 b/c incomplete
interject_yr_sex |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=prop_interject, color=as.factor(gender))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(y = "Percent Interjections", x = "Year", color = "Gender") +
  theme(legend.position="bottom")
```

@@

# plot
ggplot(interject_day_sex, aes(x=as.Date(date), y=prop_interject, color=as.factor(gender))) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber")), scales = "free_y") +
  labs(y = "Percent Interjections", x = "Date", color = "Gender") +
  theme(legend.position="bottom")
```

@@

These match well with the numbers [here](https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/sittings_statistics.pdf?la=en&hash=5B5E7E9518C4D7A416D0846FE0D44E063A4AB3FD)

@@

sitting_stats <- sitting_stats |> select(X1, X3, X5) |> 
  rename(year = X1,
         chamb = X3,
         fed_chamb = X5) |> 
  pivot_longer(chamb:fed_chamb, names_to = "fedchamb_flag", values_to = "official_stats")

# find years where there is a discrepancy in # of meetings
# care more about chamber b/c often fedchamb is wrongly transcribed so there are bound to be discrepancies
meeting_per_year |> mutate(fedchamb_flag = case_when(fedchamb_flag == 0 ~"chamb",
                                                     fedchamb_flag == 1 ~"fed_chamb"),
                           year = as.numeric(year)) |> 
  left_join(sitting_stats, by=c("year", "fedchamb_flag")) |> 
  filter(year!=2022) |> 
  mutate(diff = abs(n_sittings-official_stats)) |> 
  filter(diff!=0) |> 
  arrange(desc(diff))
```

## Q in writing and Q without notice per year

```{r}
corpus |> select(date, question, q_in_writing) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(question==1) |> 
  group_by(year, q_in_writing) |> 
  summarise(n())
```

https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/questions_statistics.pdf?la=en&hash=A3B9F79402CC7D15B576EDAF1BC28259CDEC9553

>>>>

# read in corpus
corpus <- read_parquet("/Volumes/Verbatim/v4/hansard-corpus/hansard_corpus_1998_to_2022.parquet")

@@

```{r}
### need to confirm party abbreviation mapping!!!!!
s_by_party_name <- corpus |> select(date, speech_no, party) |> 
  filter(!is.na(party)) |> 
  left_join(party_table_final, by=c("party" = "party_abb_hansard")) |> 
  group_by(date, party_abb_auspol, party_name_auspol) |> 
  summarise(n_distinct_speeches = n_distinct(speech_no, na.rm=T)) |> 
  group_by(party_abb_auspol, party_name_auspol) |> 
  summarise(total_distinct_speeches = sum(n_distinct_speeches)) |> 
  select(party_name_auspol, everything()) |> 
  arrange(desc(total_distinct_speeches)) |> 
  filter(!is.na(party_abb_auspol)) |> 
  rename("Party Name" = party_name_auspol,
         "Party Abbreviation" = party_abb_auspol,
         "Total # of Speeches" = total_distinct_speeches)

write_csv(s_by_party_name, "../paper/s_by_party_name.csv")
```

```{r}
parties_auspol <- AustralianPoliticians::get_auspol('allbyparty')

parties_auspol |> select(partyAbbrev, partyName) |> distinct()
```

@@

# plot
s_by_day |> filter(n_distinct_speeches!=0) |> 
  ggplot(aes(x=as.Date(date), y=n_distinct_speeches, color=fedchamb_flag, fill=fedchamb_flag)) + 
  geom_point(alpha=0.5) +
  geom_smooth() +
  #facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(x = "Date", y = "Number of speeches", color=NULL, fill=NULL) +
  theme_bw() +
  scale_color_discrete(labels=c("Chamber", "Federation Chamber")) +
  scale_fill_discrete(labels=c("Chamber", "Federation Chamber"))

ggsave("../paper/n_daily_speeches.png", width=8, height=5)
```

@@

# plot
# remove 2022 b/c incomplete
interject_yr_sex |> filter(year!="2022") |> 
ggplot(aes(x=as.numeric(year), y=prop_interject, color=as.factor(gender))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber"))) +
  labs(y = "Percent Interjections", x = "Year", color = "Gender") +
  theme(legend.position="bottom") +
  theme_bw()
```

@@

# plot
ggplot(interject_day_sex, aes(x=as.Date(date), y=prop_interject, color=as.factor(gender))) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess") +
  facet_wrap(~fedchamb_flag, labeller = as_labeller(c(`0` = "Chamber", `1` = "Federation Chamber")), scales = "free_y") +
  labs(y = "Percent Interjections", x = "Date", color = "Gender") +
  theme(legend.position="bottom") +
  theme_bw()
```

@@

Let's compare to the numbers [here](https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/sittings_statistics.pdf?la=en&hash=5B5E7E9518C4D7A416D0846FE0D44E063A4AB3FD).

@@

# rename select vars based on columns in pdf linked above
sitting_stats <- sitting_stats |> select(X1, X3, X5) |> 
  rename(year = X1,
         chamb = X3,
         fed_chamb = X5) |> 
  pivot_longer(chamb:fed_chamb, names_to = "fedchamb_flag", values_to = "official_stats")

# find years where there is a discrepancy in # of meetings
# care more about chamber b/c often fedchamb is wrongly transcribed so there are bound to be discrepancies
# remove 2022 b/c we don't have complete year's data
meeting_per_year |> mutate(fedchamb_flag = case_when(fedchamb_flag == 0 ~"chamb",
                                                     fedchamb_flag == 1 ~"fed_chamb"),
                           year = as.numeric(year)) |> 
  left_join(sitting_stats, by=c("year", "fedchamb_flag")) |> 
  filter(year!=2022) |> 
  mutate(diff = abs(n_sittings-official_stats)) |> 
  filter(diff!=0) |> 
  arrange(desc(diff))
```
I was able to manually check the discrepancies for the Chamber sittings, and found that there are two days in 1998 (May 6th and 7th) and two days in 2004 (February 10th and May 12th) where the House sat in the Chamber, but there is no XML transcript for Hansard available. The Federation Chamber discrepancies may be due to transcription errors - that is, Federation Chamber proceedings being nested in the Chamber proceedings node in the XML.

```{r, eval=F, echo=F}
## Q in writing and Q without notice per year
corpus |> select(date, question, q_in_writing) |>
  mutate(year = str_extract(date, "^\\d{4}")) |> 
  filter(question==1) |> 
  group_by(year, q_in_writing) |> 
  summarise(n())

#https://www.aph.gov.au/-/media/02_Parliamentary_Business/22_Chamber_Documents/224_Statistics/House_of_Representatives/Statistics_Historical/questions_statistics.pdf?la=en&hash=A3B9F79402CC7D15B576EDAF1BC28259CDEC9553
```
